---
title: "Example of Functions Part I"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vignettes/CMEvignette1.Rmd}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The goal of CME.assistant is to share reusable helper functions within the CME team.
Also feel free to download, further revise and compile locally.

```{r}
options(rmarkdown.html_vignette.check_title = FALSE)
library(CME.assistant)
```

## Functions for data engineering

### Read the final  aggregated file and output final IGME estimates in various forms
Choose output format among raw, long, wide_year and wide_ind, default to "long"  
* The long-format dataset contains __Official Name, ISO3Code, Indicator, Year, Quantile, Sex__
* The function will output all available indicators in the dataset
* Optional: subset by country ISO (`c_iso`) or year (`year_range`) 

**`get.CME.UI.data`**
```{r, eval = TRUE}
# without supplying dir_file, by default returns the published IGME 2020 result
# in long format
dt_1 <- get.CME.UI.data(use_IGME_year = 2020)
head(dt_1)
# can also output wide-format results: wide_ind, wide_year, wide_get (Rate and
# Deaths)
dt_2 <- get.CME.UI.data(use_IGME_year = 2020, format = "wide_ind")
head(dt_2)
```

```{r, eval = FALSE}
# can read final output file from Dropbox for any indicator
# loading Dropbox files requires access
dir_list <- load.final_dir()
dt_2 <- get.CME.UI.data(dir_file = dir_list$dir_female_2020)
head(dt_2)

# load all in one step
dt_all <- data.table::rbindlist(lapply(load.final_dir(2020), get.CME.UI.data, get = "both"), fill = TRUE)
dt_all[, table(Sex, Indicator)]
```


### Locate the latest master datasets directories  
Automatically point to the latest file using the date in file names  
```{r, eval = FALSE}
input_dirs <- load.IGMEinput.dir()
dir_U5MR   <- get.dir_U5MR(input_dirs$dir_IGME_20)
dir_IMR    <- get.dir_IMR(input_dirs$dir_IGME_20)
dir_NMR    <- get.dir_NMR(y5 = TRUE) # either 5-year or not
dir_gender <- get.dir_gender(plotting = TRUE) # either dataset for plotting or modeling
```


## Extra  
**`get.match`**
```{r}
# a helpful function to revise variable, you can provide a __new_list__ to define the labels
get.match(c("a","b", "c"), new_list = list("a" = "new label 1",
                                           "b" = "new label 2"))

```

**`UNICEF_colors`**
```{r}
str(UNICEF_colors)
```
