---
title: "Example of Functions Part I"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vignettes/CMEvignette1.Rmd}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The goal of CME.assistant is to share reusable helper functions within the CME team.
It is also a good idea to download, further revise and use 'Ctrl+Shift+B' to compile locally.

## UNICEF color scheme: `UNICEF_colors`
```{r}
library(CME.assistant)
str(UNICEF_colors)
```

## Functions for data engineering

### Get data in various form

Choose out format among raw, long, wide_year and wide_ind, default to "long"

**`get.CME.UI.data`**
```{r}
dt_1 <- get.CME.UI.data(c_iso = "AFG", year_range = 2015:2018, ind = "all", 
                        format = "wide_year", use_sample_data = TRUE)
dt_2 <- get.CME.UI.data(c_iso = c("AFG", "ZWE"), 
                        year_range = 2000:2018, 
                        ind = c("U5MR", "IMR"), 
                        format = "wide_ind", 
                        get = "both", 
                        use_sample_data = TRUE)
dt_3 <- get.CME.UI.data(c_iso = c("AFG", "ZWE"), 
                        year_range = 2000:2018, 
                        ind = "all", 
                        format = "wide_get", 
                        get = "both", 
                        use_sample_data = TRUE)
rmarkdown::paged_table(dt_1)
rmarkdown::paged_table(dt_2)
rmarkdown::paged_table(dt_3)
```


## Functions for making plots

**`ggsave.figs`**
```{r, fig.width = 8, fig.height = 6}
# use `ggsave.figs` to save ggplot into both pdf and png
# P.S. __SHAPforxgboost__ was my last package offering some simple ggplot visualization functions
ggplot_sample <- ggplot2::qplot(data = iris, x = Sepal.Length, y = Petal.Length)
ggsave.figs(ggplot_sample, file_name = "Iris_sample", folder_name = "figures")
# to remove the file just saved
unlink(here::here("figures"), recursive = TRUE)

# Ads: if you wish to give a try, __SHAPforxgboost__ was my last package offering some ggplot visualization functions
# install.packages("SHAPforxgboost")
SHAPforxgboost::scatter.plot.diagonal(data = iris, x = "Sepal.Length", y = "Petal.Length")
```


## Extra  
**`get.match`**
```{r}
# a helpful function to revise labels, you can provide a __new_list__ to define the labels
get.match(c("a","b", "c"), new_list = list("a" = "new label 1",
                                           "b" = "new label 2"))

```


### Data output

Maybe not that useful... put in the end. 

**`xlsx.writeMultipleData`**
```{r}
# Write several dataset into one xlsx file:
save.xlsx.XLConnect(file_dir = "temp.xlsx", list_of_dt = list(mtcars[1:5,]))
unlink("temp.xlsx")
```
