---
title: "Example of Functions Part I"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vignettes/CMEvignette1.Rmd}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The goal of CME.assistant is to share reusable helper functions within the CME team.
It is also a good idea to download, further revise and use 'Ctrl+Shift+B' to compile locally.

## UNICEF color scheme: `UNICEF_colors`
```{r}
library(CME.assistant)
str(UNICEF_colors)
```

## Functions for data engineering

### Read the final  aggregated file and output final IGME estimates in various forms
Choose output format among raw, long, wide_year and wide_ind, default to "long"  
* The long-format dataset contains __Official Name, ISO3Code, Indicator, Year, Quantile, Sex__
* The function will output all available indicators in the dataset
* Optional: subset by country ISO (`c_iso`) or year (`year_range`) 

**`get.CME.UI.data`**
```{r, eval = FALSE}
# without supplying dir_file, by default returns the published IGME 2020 final
# results
dt_1 <- get.CME.UI.data(use_IGME_year = 2020)
rmarkdown::paged_table(dt_1)

# can read final output file from Dropbox for any indicator
dir_list <- load.dir_list()
dt_2 <- get.CME.UI.data(dir_file = dir_list$dir_female_2020)
rmarkdown::paged_table(dt_2)
```


## Extra  
**`get.match`**
```{r}
# a helpful function to revise variable, you can provide a __new_list__ to define the labels
get.match(c("a","b", "c"), new_list = list("a" = "new label 1",
                                           "b" = "new label 2"))
```


### Save output
**`save.xlsx.XLConnect`**
```{r}
# Write several dataset into one xlsx file:
# save.xlsx.XLConnect(file_dir = "temp.xlsx", list_of_dt = list(mtcars[1:5,]))
# unlink("temp.xlsx")
```
